---
title: "p8105_hw1_cs3652"
author: "Chirag Shah"
date: '2018-09-18'
output: html_document
---


##Problem 1

##loading tidyverse

```{r setup, include = FALSE} 
library(tidyverse)
```

##Creating a dataframe comprising of a random sample of 10 

```{r df_problem1}
set.seed(2)

Problem1_df = tibble(
  vec_numeric = runif(10, max = 5, min = 0),
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_logical = c(vec_numeric > 2),
  vec_factor = factor(c("knicks", "raptors", "nets", "celtics", "sixers", "cavaliers", "bulls", "pacers", "pistons", "bucks"))
)
```

##Finding mean of variables

`r mean(Problem1_df$vec_numeric)`
`r mean(Problem1_df$vec_char)`
`r mean(Problem1_df$vec_logical)`
`r mean(Problem1_df$vec_factor)`

##Description of the means/results

- Finding the mean of the vec_numeric variable works because it is a numerical value and can be added and divided (which are the two operations needed to find the mean). This mean is equal to `r mean(Problem1_df$vec_numeric)`

- Finding the mean of the vec_char variable doesn't work because it is a character value which doesn't have any numerical value assigned. It is impossible to add and divide, thus when running the mean function we get `r mean(Problem1_df$vec_char)`

- Finding the mean of the vec_logical varaible gives a result. The result is the proportion of TRUE values in the entire dataset. TRUE values are given the numerical value of 1. Therefore, since there are 7 TRUE values, and 10 total values in the dataset, when running the mean function we get `r mean(Problem1_df$vec_logical)`

- Finding the mean of the vec_factor variable does not work because I have given that variable character values with no numerical values assigned. It is impossible to add and divide, thus when running the mean function we get `r mean(Problem1_df$vec_factor)`


